---
title: "Allocation Walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Allocation Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include=FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
eval = FALSE,
warning = FALSE,
error = FALSE,
message = FALSE,
echo = TRUE
)
```

```{r setup, echo=FALSE}
# remove scientific notation
options(scipen = 999)

# library packages
library(propalloc)
library(dplyr)
library(knitr)
library(kableExtra)

#' Create a custom `kable` table.
#'
#' \code{kkable} builds off of [knitr::kable()] as well
#' as [kableExtra::kable_styling()] to create a customized
#' table for use in R Markdown.
#'
#' @param data An R object, typically a matrix or data frame.
#' @param col_names column names to display - defaults to "auto" which attempts
#'   to formalize the column names for you. Alternatively, you can insert a
#'   character vector of column names.
#' @param alignment mimics [knitr::kable()] align argument but with an "auto" option.
#' @param currency_cols specify currency columns which will be formatted via [formattable::currency()].
#' @param number_cols specify numeric columns which will be formatted via [formattable::comma()].
#' @param decimal_cols specify decimal columns (i.e. numeric columns where you want to show decimal places).
#' @param decimal_digits number of digits to show in decimal cols.
#' @param percent_cols specify percent columns which will be formatted via [formattable::percent()].
#' @param percent_digits number of digits to display for percent cols.
#' @param add_totals Logical: whether to append a _totals_ row to the output. If
#'   \code{TRUE} uses \code{janitor::adorn_totals(fill = " ")} to apply
#'   the total row. See [janitor::adorn_totals()] for more information.
#' @param format_header logical - if TRUE will style the tables header.
#' @param format_totals logical - if TRUE will style the last row as a total row.
#' @param ... Other arguments passed to [knitr::kable()].
#'   (see ?knitr::kable's examples).
#' @inheritDotParams knitr::kable
#' @inheritParams kableExtra::kable_styling
#'
#' @note The tables for \code{format = 'markdown'} also work for Pandoc when the
#'   \code{pipe_tables} extension is enabled (this is the default behavior for
#'   Pandoc >= 1.10).
#'
#' @references See
#'   \url{https://github.com/yihui/knitr-examples/blob/master/091-knitr-table.Rnw}
#'    for some examples in LaTeX, but they also apply to other document formats.
#'
#' @return a customized table for display of class ""kableExtra" and/or "knitr_kable".
#'
#' @keywords internal
#'
#' @examples
#' library(propalloc)
#'
#' initial_costs <- propalloc::renewal_costs
#'
#' propalloc:::kkable(
#'  initial_costs,
#'  caption = "Initial Costs Table",
#'  proper_cols = c("cost_type", "description"),
#'  currency_cols = c("prior", "current"),
#'  digits = 0,
#'  format_header = TRUE,
#'  add_totals = TRUE,
#'  format_totals = TRUE
#' )
#'
kkable <- function(data,
                   format = "html",
                   col_names = "auto",
                   alignment = "auto",
                   proper_cols = NULL,
                   currency_cols = NULL,
                   number_cols = NULL,
                   decimal_cols = NULL,
                   decimal_digits = 2,
                   percent_cols = NULL,
                   percent_digits = 2,
                   format_header = TRUE,
                   add_totals = FALSE,
                   format_totals = add_totals,
                   digits = getOption("digits"),
                   row.names = FALSE,
                   caption = NULL,
                   label = NULL,
                   format.args = list(),
                   escape = TRUE,
                   bootstrap_options = c("striped", "bordered", "responsive"),
                   full_width = NULL,
                   position = "center",
                   font_size = NULL,
                   row_label_position = "l",
                   repeat_header_text = "\\textit{(continued)}",
                   repeat_header_method = c("append", "replace"),
                   repeat_header_continued = FALSE,
                   stripe_color = "gray!6",
                   stripe_index = NULL,
                   fixed_thead = FALSE,
                   ...) {

  if (length(col_names) == 1 && col_names == "auto") col_names <- names(data) %>% toproper()
  if (alignment == "auto") {

    classes <- get_col_classes(data)

    alignment <- purrr::map_chr(classes, function(x) {
      dplyr::case_when(
        x == "numeric" ~ "r",
        TRUE ~ "c"
      )
    }) %>%
      paste(collapse = "")

  }

  if (add_totals) {
    data <- data %>% janitor::adorn_totals(fill = " ")
  }

  if (!is.null(proper_cols)) {
    # data <- data %>% dplyr::mutate_at(dplyr::vars(tidyselect::all_of(proper_cols)), toproper)
    data <- data %>% dplyr::mutate_at(dplyr::vars(proper_cols), toproper)
  }

  if (!is.null(currency_cols)) {
    # data <- data %>% dplyr::mutate_at(dplyr::vars(tidyselect::all_of(currency_cols)), formattable::currency, digits = 0)
    data <- data %>% dplyr::mutate_at(dplyr::vars(currency_cols), formattable::currency, digits = 0)
  }

  if (!is.null(number_cols)) {
    # data <- data %>% dplyr::mutate_at(dplyr::vars(tidyselect::all_of(number_cols)), formattable::comma, digits = 0)
    data <- data %>% dplyr::mutate_at(dplyr::vars(number_cols), formattable::comma, digits = 0)
  }

  if (!is.null(percent_cols)) {
    # data <- data %>% dplyr::mutate_at(dplyr::vars(tidyselect::all_of(percent_cols)), formattable::percent, digits = percent_digits)
    data <- data %>% dplyr::mutate_at(dplyr::vars(percent_cols), formattable::percent, digits = percent_digits)
  }

  if (!is.null(decimal_cols)) {
    # data <- data %>% dplyr::mutate_at(dplyr::vars(tidyselect::all_of(decimal_cols)), formattable::comma, digits = decimal_digits)
    data <- data %>% dplyr::mutate_at(dplyr::vars(decimal_cols), formattable::comma, digits = decimal_digits)
  }

  hold <- knitr::kable(data,
                       format = format,
                       digits = digits,
                       row.names = row.names,
                       col.names = col_names,
                       align = alignment,
                       caption = caption,
                       label = label,
                       format.args = format.args,
                       escape = escape
  ) %>%
    kableExtra::kable_styling(position = position,
                              bootstrap_options = bootstrap_options,
                              full_width = full_width,
                              font_size = font_size,
                              row_label_position = row_label_position,
                              repeat_header_text = repeat_header_text,
                              repeat_header_method = repeat_header_method,
                              repeat_header_continued = repeat_header_continued,
                              stripe_color = stripe_color,
                              stripe_index = stripe_index,
                              fixed_thead = fixed_thead)

  if (format_header) hold <- hold %>% kableExtra::row_spec(row = 0, bold = TRUE, align = "center")
  if (format_totals) hold <- hold %>% kableExtra::row_spec(row = nrow(data), bold = TRUE, background = "lightgrey")

  return(hold)

}
```

## Overview

The purpose of this report is to walk through the various allocation procedures 
utilized in the shiny application. 

In order to provide knowledgeable insights from this report, we have supplied 
reasonable inputs that mimic the *user-driven* aspects of the allocation 
procedures. 

## Setup

The first step is to ingest all of the various user-driven inputs and demo-data to form a dataset to perform the allocation with. 

```{r load_demo_data}
load_demo_data()
```

Specify various 'user-defined' inputs: 

- Experience Period for Loss Run 
- Capping Threshold (%)
- Budget guidance increase factor (a % applied to prior rates which are then used to allocate premiums)

```{r user-inputs}
# specify experience period based on loss run
min_year <- min(loss_run$year)
max_year <- max(loss_run$year)
experience_period <- c(min_year:max_year)

experience_period_display <- paste0(
  lubridate::ymd(paste0(min_year, "-01-01")) %>% format("%B %d, %Y"), " to ",
  lubridate::ymd(paste0(max_year, "-12-31")) %>% format("%B %d, %Y")
)

# specify percent change capping threshold
cap_threshold <- .25

# A 5% budget guidance increase factor
budget_guidance_percent <- 0.05
```

### Costs to be Allocated

we need to extract the ***costs to be allocated*** in the model, specifically: 

1. The Total Cost Including Expenses  
1. The Risk Transfer Cost  
1. The All Risk Cost  
1. The Sum of Expenses  

To do this, I utilize an internal, custom utility function `extract_costs_for_allocation()`.

The resulting ***Costs for Allocation*** are:

```{r extract_costs}
# extract costs from renewal cost table
costs <- extract_costs(renewal_costs)

# derive current and prior overall rates and percent change
curr_rate <- costs$risk_transfer / sum(sov$tiv)
prior_rate <- sum(priors$prior_risk_transfer_premium) /
  sum(priors$prior_tiv)
pct_change <- (curr_rate / prior_rate) - 1

tibble::as_tibble(costs) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "Cost Type") %>%
  dplyr::rename("Dollar Amount" = V1) %>%
  tibble::as_tibble() %>%
  kkable(currency_cols = "Dollar Amount",
         proper_cols = "Cost Type",
         caption = "Extracted Costs for Allocation:")
```


### Derive Relativity Adjusted TIVs 

This step involves taking the user-defined relativity's and multiplying them by 
the initial *Total Insured Values* defined in the Schedule of Values (SOV). 

Specifically, we take the user-input *Relativity Table* and apply a custom utility 
function `derive_relativity_adjusted_tivs()` to derive our final relativity 
adjusted TIVs. 

First, lets look at the input relativities: 

```{r preview_relativities, results='asis'}
rel_tables <- list(bu_rels, combustible_rels, sprinkler_tier_rels) %>%
  purrr::set_names(c("Business Unit Relativity",
                     "AOP Combustible Relativity",
                     "AOP Sprinkler Tier Relativity")) %>%
  purrr::map(apply_labels, dict = dictionary, dataset_name = "rels") %>%
  purrr::map(tibble::as_tibble)

purrr::map2(rel_tables, names(rel_tables), function(x, y) {
  kkable(data = x, caption = y, col_names = NULL, digits = 3, add_digits = TRUE)
})
```

### Apply relativity factors to TIVs

Relativities are then applied directly to each entity's TIV. For each
relativity type, a separate relativity-adjusted TIV is calculated and
used as required in eithe catastrophy allocation, All Other Peril allocation,
or terrorism allocation.

```{r rel_adj, results='asis'}
rels_list <- list(
    relativity_data = list(
      bu_rels[, c(1, 2)], bu_rels[, c(1, 3)], bu_rels[, c(1, 4)], bu_rels[, c(1, 5)],
      bu_rels[, c(1, 6)], sprinkler_tier_rels, combustible_rels
    ),
    coverage = list(
      "aop", "cat_eq", "cat_wind", "cat_flood", "terrorism", "aop", "aop"
    ),
    sov_linker = list(
      "bu", "bu", "bu", "bu", "bu", "aop_sprinkler_tier", "aop_combustible"
    )
  )

rel_adjusted_tivs <- ingest_relativities(rels_list, sov = sov)

kkable(head(rel_adjusted_tivs),
       proper_cols = "entity_id", 
       currency_cols = names(rel_adjusted_tivs[2:ncol(rel_adjusted_tivs)]))
```

### Prepare claim surcharges

For entities that have experienced claims, the client may wish to increase
their allocated premium. The rules for doing so are limited in this application
to: applying a % surcharge to the TIV for each claim made by the entity in a
certain time period, or adding a $ surcharge to the allocated premium (at this
stage in the calculation) for each claim made in a certain time period.

The demo data, `count_buckets`, sets out a typical specification for such
functionality.

```{r count_buckets, results='asis'}
count_buckets %>%
  select(name:dollar_surcharge) %>%
  kkable(
    col_names = c("Label", "Minimum", "Maximum", "Percent Surcharge", "Dollar Surcharge"),
    caption = "User-Defined Claim Count Bucket Surcharges",
    currency_cols = c("min", "max", "dollar_surcharge"),
    percent_cols = c("percent_surcharge")
  )
```

Here, `count_buckets` are applied to the loss run ready for surcharging
the premiums calculated in the main part of the model:

```{r derive_count_buckets_by_entity}
entity_loss_data <- entity_loss_summary(loss_run, count_buckets, experience_period)

col_names <- c("Entity ID", 
               count_buckets$name,
               "Total Counts", "Total Incurred")

kkable(entity_loss_data %>%
         # dplyr::mutate(entity_id = toproper(entity_id)) %>%
         dplyr::arrange(dplyr::desc(total_incurred)) %>%
         head(10), 
       col_names = col_names,
       proper_cols = "entity_id",
       currency_cols = length(col_names),
       caption = "Summarized Loss Data by Entity (Top 10 Entities by Total Incurred)")
```

### Putting it all together

The steps involved are:

1. `merge_entity_data()`:
  - Apply `rates` to `relativity_adjusted_tivs` and
    also to `priors` (to simultaneously obtain an allocation of
    prior premiums)
  - Apply `entity_loss_data` to surcharge the resulting premiums
    pased on `count_buckets` and the loss run experience

2. `preliminary_allocation()`:
  - Determine a preliminary allocation of premiums (post claim surcharges)
    by calibrating the total allocated premium to match those indicated
    in `costs`
  - Also, apply `budget_guidance_percent` to prior premium rates
    to perform this allocation scenario
    
3. `apply_threshold()`:
  - Take the preliminary allocation of premiums and enforce a threshold upon
    them which prevents premiums allocated in the current period deviating
    too far from premium allocations of the prior period. For example, if the
    threshold is set to 25% (`threshold <- 0.25`), and if the prior
    premium allocated for a particular entity was 1m USD, then roughly
    the current premium allocation would be between 750k USD, and 1.25m USD.
    
4. Perform final calculations to include expenses in the allocation.
    

```{r allocate}
# merge entity data (sov, rel adjusted tivs, loss data, market and model rates,
# and priors)

allocation_data <- merge_entity_data(
  sov,
  rel_adjusted_tivs,
  entity_loss_data,
  rates,
  priors
) %>%
  # perform initial preliminary allocation (CAT first, back into AOP, terror)
  # this is uncapped. before surcharges, and excluding expenses
  preliminary_allocation(costs, budget_guidance_percent) %>%
  # apply surcharges
  apply_surcharges(count_buckets) %>%
  # adjust column names for apply threshold function
  # TODO: add arguments to apply threshold functions for specifying column
  # so don't have to add this step.
  mutate(
    prior_allocated = prior_risk_transfer_premium,
    prior_allocated_rate = prior_allocated / prior_tiv,
    uncapped_allocated = surcharged_premium
  ) %>%
  # apply capping using a default 25% threshold
  # TODO: add argument to apply threshold for whether or not to net the
  # total pct change or not - currently it does this
  apply_threshold(
    total_pct_chg = pct_change,
    threshold = cap_threshold
  ) %>%
  # final rebalancing and allocate expenses
  allocate_expenses(costs, weight_variable = "tiv")

```

And the result is:

```{r cleanup, results='asis'}
# declutter results and output
allocation_data %>%
  mutate(current_allocated_rate = rebalanced_allocated / tiv) %>% # excl expense
  select(
    entity_id,
    tiv,
    prior_tiv,
    aop_adj_tiv:terrorism_adj_tiv,
    model_aop_rate,
    model_cat_eq_rate,
    model_cat_wind_rate,
    model_cat_flood_rate,
    model_terrorism_rate,
    prior_risk_transfer_premium,
    preliminary_model_premium = total_model_premium_adj,
    surcharge,
    surcharged_premium,
    prior_allocated,
    prior_allocated_rate,
    uncapped_allocated,
    capped_allocated = allocated,
    final_allocated = rebalanced_allocated,
    current_allocated_rate,
    rate_percent_change = capped_rate_percent_change,
    allocated_expenses,
    final_allocated_w_expense
  ) %>%
  kkable()
```


### Driver Summary

`propalloc` comes with functionality to split out the effects of
each of the stages of premium allocation as they contribute to its change
between prior and current premium allocations. This function uses argument
`filter_vector` to show such a split for a subset of entities if desired.

```{r driver_summary}
driver_summary <- prepare_driver_summary(allocation_data, filter_vector = c())
driver_summary_bu_b_only <- prepare_driver_summary(allocation_data, filter_vector = c(bu = "bu_b"))
```
