---
title: "Allocation Walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Allocation Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include=FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
eval = FALSE,
warning = FALSE,
error = FALSE,
message = FALSE,
echo = TRUE
)
```

```{r setup, eval=TRUE, echo=FALSE}
# remove scientific notation
options(scipen = 999)

# library packages
library(propalloc)
library(dplyr)
```

## Overview

The purpose of this report is to walk through the various allocation procedures 
utilized in the shiny application. 

In order to provide knowledgeable insights from this report, we have supplied 
reasonable inputs that mimic the *user-driven* aspects of the allocation 
procedures. 

## Setup

The first step is to ingest all of the various user-driven inputs and demo-data to form a dataset to perform the allocation with. 

```{r load_demo_data, eval=TRUE}
load_demo_data()
dictionary <- propalloc:::dictionary
```

Specify various 'user-defined' inputs: 

- Experience Period for Loss Run 
- Capping Threshold (%) 

```{r user-inputs, eval=TRUE}
# specify experience period based on loss run
min_year <- min(loss_run$year)
max_year <- max(loss_run$year)
experience_period <- c(min_year:max_year)

experience_period_display <- paste0(
  lubridate::ymd(paste0(min_year, "-01-01")) %>% format("%B %d, %Y"), " to ",
  lubridate::ymd(paste0(max_year, "-12-31")) %>% format("%B %d, %Y")
)

# specify percent change capping threshold
cap_threshold <- .25
```

### Costs to be Allocated

we need to extract the ***costs to be allocated*** in the model, specifically: 

1. The Total Cost Including Expenses  
1. The Risk Transfer Cost  
1. The All Risk Cost  
1. The Sum of Expenses  

To do this, I utilize an internal, custom utility function \code{extract_costs_for_allocation()}.  

The resulting ***Costs for Allocation*** are:

```{r extract_costs, eval=TRUE}
# extract costs from renewal cost table
costs <- extract_costs(renewal_costs)

# derive current and prior overall rates and percent change
curr_rate <- costs$risk_transfer / sum(sov$tiv)
prior_rate <- sum(priors$prior_risk_transfer_premium) /
  sum(priors$prior_tiv)
pct_change <- (curr_rate / prior_rate) - 1

tibble::as_tibble(costs) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "Cost Type") %>%
  dplyr::rename("Dollar Amount" = V1) %>%
  tibble::as_tibble() %>%
  kkable(currency_cols = "Dollar Amount",
         proper_cols = "Cost Type",
         caption = "Extracted Costs for Allocation:")
```


### Derive Relativity Adjusted TIVs 

This step involves taking the user-defined relativity's and multiplying them by 
the initial *Total Insured Values* defined in the Schedule of Values (SOV). 

Specifically, we take the user-input *Relativity Table* and apply a custom utility 
function `derive_relativity_adjusted_tivs()` to derive our final relativity 
adjusted TIVs. 

First, lets look at the input relativities: 

```{r preview_relativities, eval=TRUE, results='asis'}
rel_tables <- list(bu_rels, combustible_rels, sprinkler_tier_rels) %>%
  purrr::set_names(c("Business Unit Relativity",
                     "AOP Combustible Relativity",
                     "AOP Sprinkler Tier Relativity")) %>%
  purrr::map(apply_labels, dict = dictionary, dataset_name = "rels") %>%
  purrr::map(tibble::as_tibble)

purrr::map2(rel_tables, names(rel_tables), function(x, y) {
  kkable(data = x, caption = y, col_names = NULL, digits = 3, add_digits = TRUE)
})
```

```{r rel_adj, eval=TRUE, results='asis'}
rel_adjusted_tivs <- apply_rels()

kkable(head(rel_adjusted_tivs),
       proper_cols = "entity_id", 
       currency_cols = names(rel_adjusted_tivs[2:ncol(rel_adjusted_tivs)]))
```

```{r count_buckets, results='asis', eval=TRUE}
count_buckets %>%
  select(name:dollar_surcharge) %>%
  kkable(
    col_names = c("Label", "Minimum", "Maximum", "Percent Surcharge", "Dollar Surcharge"),
    caption = "User-Defined Claim Count Bucket Surcharges",
    currency_cols = c("min", "max", "dollar_surcharge"),
    percent_cols = c("percent_surcharge")
  )
```

```{r derive_count_buckets_by_entity, eval=TRUE}
entity_loss_data <- entity_loss_summary(loss_run, count_buckets, experience_period)

col_names <- c("Entity ID", 
               count_buckets$name,
               "Total Counts", "Total Incurred")

kkable(entity_loss_data %>%
         # dplyr::mutate(entity_id = toproper(entity_id)) %>%
         dplyr::arrange(dplyr::desc(total_incurred)) %>%
         head(10), 
       col_names = col_names,
       proper_cols = "entity_id",
       currency_cols = length(col_names),
       caption = "Summarized Loss Data by Entity (Top 10 Entities by Total Incurred)")
```

```{r allocate, eval=TRUE}
# merge entity data (sov, rel adjusted tivs, loss data, market and model rates, and priors)
allocation_data <- merge_entity_data(sov,
                                     rel_adjusted_tivs,
                                     entity_loss_data,
                                     rates,
                                     priors) %>%
  # perform initial preliminary allocation (CAT first, back into AOP, terror)
  # this is uncapped. before surcharges, and excluding expenses
  preliminary_allocation() %>%
  # apply surcharges
  apply_surcharges(count_buckets) %>%
  # adjust column names for apply threshold function
  # TODO: add arguments to apply threshold functions for specifying column
  # so don't have to add this step.
  mutate(prior_allocated = prior_risk_transfer_premium,
         prior_allocated_rate = prior_allocated / prior_tiv,
         uncapped_allocated = surcharged_premium) %>%
  # apply capping using a default 25% threshold
  # TODO: add argument to apply threshold for whether or not to net the
  # total pct change or not - currently it does this
  apply_threshold(total_pct_chg = pct_change,
                  threshold = cap_threshold) %>%
  # final re-balancing and allocate expenses
  mutate(
    # coalesce no longer necessary since removed "new" entities
    # allocated = coalesce(allocated, surcharged_premium),
    percent_tiv = tiv / sum(tiv),
    allocated_pct = allocated / sum(allocated, na.rm = TRUE),
    rebalanced_allocated = allocated_pct * costs$risk_transfer,
    allocated_expenses = percent_tiv * costs[["expenses"]],
    final_allocated_w_expense = rebalanced_allocated + allocated_expense,
    # derive rate (excluding expense)
    current_allocated_rate = rebalanced_allocated / tiv,
    # create cols for risk-engineering and man_adj
    risk_engineering = 0, # TODO: need to mimic user-input
    # final rebalance
    final_allocated_w_risk_engineering = final_allocated_w_expense + 
      risk_engineering,
    final_allocated_w_risk_engineering = (final_allocated_w_expense / 
                                            sum(.data$final_allocated_w_expense)) * 
      sum(.data$final_allocated_w_expense),
    manual_adjustments = 0, # TODO: need to mimic user-input
    final_allocated_w_man_adj = final_allocated_w_risk_engineering + 
      manual_adjustments
  )

# validate results and output ---------------------------------------------

# checks
all(
  sum(allocation_data$rebalanced_model_terrorism_premium_adj) == costs$terrorism,
  sum(allocation_data$total_model_premium_adj) == costs$risk_transfer,
  sum(allocation_data$rebalanced_model_aop_premium_adj, allocation_data$total_model_cat_premium_adj) == costs$all_risk,
  sum(allocation_data$final_allocated_w_expense) == costs$risk_transfer + costs$expenses,
  sum(allocation_data$final_allocated_w_expense) == costs$total_incl_expense,
  sum(allocation_data$final_allocated_w_risk_engineering) == costs$total_incl_expense
)

# declutter results and output
allocation_results <- allocation_data %>%
  select(
    entity_id,
    tiv,
    prior_tiv,
    aop_adj_tiv:terrorism_adj_tiv,
    model_aop_rate,
    model_cat_eq_rate,
    model_cat_wind_rate,
    model_cat_flood_rate,
    model_terrorism_rate,
    prior_risk_transfer_premium,
    preliminary_model_premium = total_model_premium_adj,
    surcharge,
    surcharged_premium,
    prior_allocated,
    prior_allocated_rate,
    uncapped_allocated,
    capped_allocated = allocated,
    final_allocated = rebalanced_allocated,
    current_allocated_rate,
    rate_percent_change = capped_rate_percent_change,
    allocated_expense,
    final_allocated_w_expense
  )
```

Outputs to Excel for review: 

```{r outputs, eval=FALSE}
inputs <- list(
  "Dictionary" = dictionary,
  "SOV" = sov,
  "Renewal Costs" = renewal_costs,
  "Priors" = priors,
  "Loss Run" = loss_run,
  "Count Buckets" = count_buckets,
  "Rates" = rates,
  "BU Relativity" = bu_rels,
  "Combustible Relativity" = combustible_rels,
  "Sprinkler Relativity" = sprinkler_tier_rels
)

outputs <- list(
  "Full Allocation Data" = allocation_data,
  "Simple Allocation Results" = allocation_results
)

writexl::write_xlsx(inputs, "../outputs/2020-04-15-allocation-inputs.xlsx")
writexl::write_xlsx(outputs, "../outputs/2020-04-15-allocation-results.xlsx")
```






